\myTitle

### 2024-06-28 tex/abstract.tex
Minor edits

We need to write an abstract.

>>>>

Genetic methods have become an essential component of ecological investigation
and conservation planning for fish and wildlife. Among these methods is the use
of genetic marker data to identify individuals to
populations, or stocks, of origin. More recently, methods that involve genetic
pedigree reconstruction to identify relationships between individuals within populations
have become common. We present, here, a novel set of
multi-allelic microhaplotype genetic markers for Chinook salmon which provide unprecedented
resolution for population discrimination and relationship identification for a rapidly and
economically assayed panel of markers. We show how this set of microhaplotypes
provides definitive power to identify all known lineages of Chinook salmon in
California. The inclusion of genetic loci that have known associations with phenotype
and that were identified as outliers in examination of whole genome sequence data,
allows resolution of stocks that are not highly genetically differentiated but
are phenotypically distinct and managed as such. 
This same set of multiallelic genetic markers have ample variation to
accurately identify parent-offspring and full-sibling
pairs in all California populations, including the genetically depauperate winter-run
lineage.
Validation of this marker panel in salmon populations not previously studied with
modern genetic methods, also reveals novel biological insights, including
the presence of a single copy of a haplotype for a phenotype that has not
been documented in that part of the species range, and a clear signal of mixed
ancestry for a salmon population that is on the geographic margins of the 
primary evolutionary lineages present in California.

### 2024-06-28 tex/author-title-etc.tex
More overleaf changes
Minor edits

%%%%% THIS IS THE SECTION WHERE THE AUTHOR PUTS IN ALL OF THEIR TITLE AND AFFILIATION
%%%%% INFORMATION AND  A FEW OTHER THINGS
\newcommand{\myTitle}{A multipurpose microhaplotype panel for analysis of California Chinook salmon}
\title{\myTitle}

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{Eric C. Anderson$^{*,\S}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
\author{Eric C. Anderson$^{*}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{
$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA, Santa Cruz, California, USA. $^\dagger$Current address: Mac's home address in AK.
}

>>>>

%%%%% THIS IS THE SECTION WHERE THE AUTHOR PUTS IN ALL OF THEIR TITLE AND AFFILIATION
%%%%% INFORMATION AND  A FEW OTHER THINGS
\newcommand{\myTitle}{A multipurpose microhaplotype panel for genetic analysis of California Chinook salmon}
\title{\myTitle}

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{Eric C. Anderson$^{*,\S}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, Neil F. Thompson$^{*,\ddag}$, John Carlos Garza$^{*}$}
\author{Eric C. Anderson$^{*}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, Neil F. Thompson$^{*,\ddag}$, John Carlos Garza$^{*}$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{
$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA, Santa Cruz, California, USA. $^\dagger$Current address: Mac's home address in AK. $^\ddag$Current address: Get Neil's address.
}

### 2024-06-28 tex/body-text-of-paper.tex
Minor edits

The emergence over the last decade of parentage based tagging (PBT) as a management
tool for Pacific salmon adds another factor to consider when developing a GSI baseline.
Since the first proposal \citep{anderson2005description} to replace or augment the coded-wire tag
programme \citep{nandor2010overview}
with PBT, it has been noted that one of the major advantages of a genetic program for
PBT is that the genetic markers used for PBT could also be useful for GSI (and
vice versa).

Thus, any panel of markers to be used for GSI (or PBT) should also be evaluated on its
utility for PBT (or GSI).  PBT has been remarkably successful in fisheries
management, having been used for over a decade in the management of Chinook salmon
and steelhead trout in major basins of the Columbia River \cite{steele2019parentage,horn2023multigeneration}, and having been employed to dramatically further our
understanding of the genetic inheritance of key traits in salmonids
\citep{abadia2013large,beulke2023distinct}. However, PBT is just one subset of a whole
family of statistical genetic methods employing relationship inference to learn about
populations.  For example inference of the full siblings amongst a sample of fish
provides information about the effective number of adults producing offspring
\citep{waples2011inbreeding,wang2023estimating}, which can
be a valuable source of information when sampling of the adults is not possible.

>>>>

Over the last decade an additional genetic method, parentage based tagging, or PBT,
\citep{anderson2005description,steele2019parentage}
has become established as an extremely valuable management
tool for Pacific salmon.  The availability of such family-based methods adds another factor to
consider when developing a GSI baseline.
Since the first proposal \citep{anderson2005description} to replace or augment the coded-wire tag
programme \citep{nandor2010overview}
with PBT, it has been noted that one of the major advantages of a genetic program for
PBT is that the genetic markers used for PBT could also be useful for GSI (and
vice versa).  
Thus, any panel of markers to be used for GSI (or PBT) should also be evaluated on its
utility for PBT (or GSI).

PBT has been remarkably successful in fisheries
management, having been used for over a decade in the management of Chinook salmon
and steelhead trout in major basins of the Columbia River \cite{steele2019parentage,horn2023multigeneration}, and having been employed to dramatically further our
understanding of the genetic inheritance of key traits in salmonids
\citep{abadia2013large,beulke2023distinct}. However, PBT is just one subset of a whole
family of statistical genetic methods employing relationship inference to learn about
populations.  For example inference of the full siblings amongst a sample of fish
provides information about the effective number of adults producing offspring
\citep{waples2011inbreeding,wang2023estimating}, which can
be a valuable source of information when sampling of the adults is not possible.
